---
title: "hNPC_scRNAseq_analysis"
author: "SabaShahin"
date: "06/06/2023"
output: html_document
    
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


load required libraries
```{r}
suppressWarnings(library(dplyr))
suppressWarnings(library(Seurat))
suppressWarnings(library(tidyverse))
suppressWarnings(library(Matrix))
suppressWarnings(library(ggplot2))
suppressWarnings(library(cowplot))
suppressWarnings(library(units))
suppressWarnings(library(umap))
suppressWarnings(library(cluster))
suppressWarnings(library(tsne))
suppressWarnings(library(viridis))
suppressWarnings(library(patchwork))
suppressWarnings(library(pmap))
suppressWarnings(library(irlba))
suppressWarnings(library(beepr))
suppressWarnings(library(gridExtra))
suppressWarnings(library(Rcpp))
suppressWarnings(library(png))
suppressWarnings(library(harmony))
```

# set working directory (main directory in which all data sets are present)
# get data location in the working directory (folder that has 10X data/filtered_feature_bc_matrix data)

Make sure you have all three file in the correct directory 

1. matrix.mtx
2. genes.tsv
3. barcodes.tsv

```{r}

setwd("~/Desktop/invitro_invivo_hNPC_snRNAseq_data")
dirs <- list.dirs(path = 'hNPC_scRNAseq_data', recursive = F, full.names = F)

```

# load data 
# Create each individual Seurat object for every sample


```{r}

for(x in dirs){
  name <- gsub('_filtered_feature_bc_matrix','', x)
  
  cts <- ReadMtx(mtx = paste0('hNPC_scRNAseq_data/',x,'/matrix.mtx.gz'),
                 features = paste0('hNPC_scRNAseq_data/',x,'/features.tsv.gz'),
                 cells = paste0('hNPC_scRNAseq_data/',x,'/barcodes.tsv.gz'))
  
  assign(name, CreateSeuratObject(counts = cts, min.cells = 3, min.features = 200))
}

```
# check object in environment, type ls() in console
# create new directory to save the figures
# Set working dir 
# save session info

```{r}

# check the list in console
ls()
setwd("~/Desktop/08-10-2023_hNPC_scRNA_seq_analysis")
```

# save session info

```{r}
sessionInfo()

date <- 08-10-2023
project <- "hNPC"

sink(paste0(date,"_",project,"_devtools_sessionInfo.txt"))
devtools::session_info()
sink()


sink(paste0(date,"_",project,"_sessionInfo.txt"))
sessionInfo()
sink()

```


# create merge seurat objects
# merge datasets
```{r}

hNPCs_merged <- merge(invitro_hNPCsample14, y = c(invitro_hNPCsample15, invitro_hNPCsample16, invivoP60_hNPCsampleA, invivoP60_hNPCsampleB, invivoP90_hNPCsampleD),
                      add.cell.ids = c("invitro_hNPCsample14","invitro_hNPCsample15", "invitro_hNPCsample16", "invivoP60_hNPCsampleA", "invivoP60_hNPCsampleB", "invivoP90_hNPCsampleD"),
                      project = 'hNPCs')


## save object
save(hNPCs_merged, file = "hNPCs_merged.rds")

# remove all objects that will not be used
rm(invitro_hNPCsample14, invitro_hNPCsample15, invitro_hNPCsample16, invivoP60_hNPCsampleA, invivoP60_hNPCsampleB, invivoP90_hNPCsampleD)

# run garbage collect to free up memory
gc()

# check no. of cells and features in merged seurat object
hNPCs_merged

# see the count matrix and the metatada for few cells
as.data.frame(hNPCs_merged@assays$RNA@counts[1:10, 1:2])
head(hNPCs_merged@meta.data, 10)
tail(hNPCs_merged@meta.data, 10)

```


# create a sample column and split sample column
```{r}
hNPCs_merged$sample <- rownames(hNPCs_merged@meta.data)

View(hNPCs_merged@meta.data)

# split sample column

hNPCs_merged@meta.data <- separate(hNPCs_merged@meta.data, col = 'sample', into = c('Condition',"Sample_ID",'Barcode'),
                                   sep = '_')
# View metadata
View(hNPCs_merged@meta.data)

# lets check the no. of cells in each conditions
table(hNPCs_merged$Condition)
```

sanity check (whether all group are present in the data or not), type unique(hNPCs_merged_seurat@meta.data$Condition) in console

# Before QC & filtering -----------------------

## calculate some additional metrics for plotting:

1. number of genes detected per UMI: this metric with give us an idea of the complexity of our dataset (more genes detected per UMI, more complex our data

# Add number of genes per UMI for each cell to metadata
```{r}
hNPCs_merged$log10GenesPerUMI <- log10(hNPCs_merged$nFeature_RNA) / log10(hNPCs_merged$nCount_RNA)

# View metadata
View(hNPCs_merged@meta.data)
```

2. Check if seurat object contain mitochondrial gene and and column for mitoPercent in metadata
```{r}
# Check if seurat object contain mitochondrial gene
str_subset(rownames(hNPCs_merged), "^MT")

# calculate the percentage of mitochondrial reads per cell and column for mitoPercent
hNPCs_merged$mitoPercent <- PercentageFeatureSet(hNPCs_merged, pattern = "^MT-")
# this code also add column for mitoPercent hNPCs_merged_seurat[[percent.mt]] <- PercentageFeatureSet(hNPCs_merged_seurat, pattern = "^MT-")

## View metadata
View(hNPCs_merged@meta.data)
```

3. calculate the  proportion gene expression that comes from ribosomal proteins and add column for riboPercent in metadata
```{r}

hNPCs_merged$riboPercent <- PercentageFeatureSet(hNPCs_merged, pattern = "^RP[SL]")

# View metadata
View(hNPCs_merged@meta.data)
```


# Visualize QC metrics as a violin plot
```{r}
# plot distributions of QC metrics, grouped by SampleID with 
pdf('ViolinPlot_before_QC_with_Points.pdf',height=6,width=12)
VlnPlot(hNPCs_merged, features = c("nFeature_RNA", "nCount_RNA", "mitoPercent", "riboPercent"), ncol = 4, pt.size=.2, group.by='Condition')
dev.off()


# as a violin plot
pdf('ViolinPlot_before_QC_no_Points.pdf', width=4, height=18)
VlnPlot(
  hNPCs_merged,
  features = c("nFeature_RNA", "nCount_RNA", "mitoPercent", "riboPercent"),
  group.by='Condition',
  ncol = 1, pt.size=0)
dev.off()

```


# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
```{r}
pdf('FeatureScatterPlot_before_filtering.pdf',height=5,width=15)
plot1 <- FeatureScatter(hNPCs_merged, feature1 = "nCount_RNA", feature2 = "mitoPercent", group.by = "Condition")
plot2 <- FeatureScatter(hNPCs_merged, feature1 = "nCount_RNA", feature2 = "riboPercent", group.by = "Condition")
plot3 <- FeatureScatter(hNPCs_merged, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "Condition") + geom_smooth(method = lm)
plot1 + plot2 + plot3
dev.off()
```



#Let’s get a sense for the size of our dataset by plotting the number of cells in each sample.
```{r}
theme_set(theme_cowplot())

# use table function to get the number of cells in each Sample as a dataframe
df <- as.data.frame(rev(table(hNPCs_merged$Condition)))
colnames(df) <- c('Condition', 'n_cells')


# bar plot of the number of cells in each sample
p <- ggplot(df, aes(y=n_cells, x=reorder(Condition, -n_cells), fill=Condition)) +
  geom_bar(stat='identity') +
  scale_y_continuous(expand = c(0,0)) +
  NoLegend() + RotatedAxis() +
  ylab(expression(italic(N)[cells])) + xlab('Condition') +
  ggtitle(paste('Pre_filtering:', sum(df$n_cells))) +
  theme(
    panel.grid.minor=element_blank(),
    panel.grid.major.y=element_line(colour="lightgray", size=0.5),
  )

png('Total_cells_prefiltering.png', width=3, height=4, res=200, units='in')
print(p)
dev.off()

```

# Visualize the number UMIs/transcripts per cell color=Condition,
```{r}
# Create metadata dataframe
metadata <- hNPCs_merged@meta.data

p1 <- metadata %>% ggplot(aes(x=nCount_RNA, fill= Condition)) + 
  geom_density(alpha = 0.2) + 
  scale_x_log10() + 
  theme_classic() +
  ylab("cell density") +
  geom_vline(xintercept = 500)

png('nCount_per_cell.png', width=6, height=4, res=200, units='in')
print(p1)
dev.off()

```


# Visualize the distribution of genes detected per cell via histogram
```{r}

p2 <- metadata %>% 
  ggplot(aes(color=Condition, x=nFeature_RNA, fill= Condition)) + 
  geom_density(alpha = 0.2) + 
  theme_classic() +
  scale_x_log10() + 
  geom_vline(xintercept = 300)

png('Genes_per_cell.png', width=6, height=4, res=200, units='in')
print(p2)
dev.off()

```

# Visualize the distribution of genes detected per cell via boxplot

```{r}
p3 <- metadata %>% 
  ggplot(aes(x=Condition, y=log10(nFeature_RNA), fill=Condition)) + 
  geom_boxplot() + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("NCells vs NGenes")

png('Genes_per_cell_boxplot.png', width=4, height=4, res=200, units='in')
print(p3)
dev.off()

```

## UMIs vs. genes detected
# Visualize the correlation between genes detected and number of UMIs and determine whether strong presence of cells with low numbers of genes/UMIs

```{r}
p4 <- metadata %>% 
  ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=mitoPercent)) + 
  geom_point() + 
  scale_colour_gradient(low = "gray90", high = "black") +
  stat_smooth(method=lm) +
  scale_x_log10() + 
  scale_y_log10() + 
  theme_classic() +
  geom_vline(xintercept = 500) +
  geom_hline(yintercept = 300) +
  facet_wrap(~Condition)

png('Corellation_plot.png', width=10, height=4, res=200, units='in')
print(p4)
dev.off()

```

# Visualize the distribution of mitochondrial gene expression detected per cell
```{r}
p5 <- metadata %>% 
  ggplot(aes(color=Condition, x=mitoPercent, fill=Condition)) + 
  geom_density(alpha = 0.2) + 
  scale_x_log10() + 
  theme_classic() +
  geom_vline(xintercept = 0.2)

png('MitoGene_Cell.png', width=10, height=4, res=200, units='in')
print(p5)
dev.off()

```

# Visualize the overall complexity of the gene expression by visualizing the genes detected per UMI

```{r}
p6 <- metadata %>%
  ggplot(aes(x=log10GenesPerUMI, color = Condition, fill=Condition)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  geom_vline(xintercept = 0.8)

png('GeneExpressionComplexity.png', width=6, height=4, res=200, units='in')
print(p6)
dev.off()

```

# save the merged dataset
```{r}
saveRDS(hNPCs_merged, file = "hNPCs_merged.rds")

```

# assign data to new object
# create new directory to save the figures
```{r}

hNPCs_Filtered <- hNPCs_merged

##  create new directory to save the figures
dir.create("08102023_hNPC_Filtered")
setwd("08102023_hNPC_Filtered")

# remove unused dataset
rm(hNPCs_merged)
```

# QC filtering -----------------------
Final filtering, change object name as with same object name it will overwrite, object name= merged_seurat_filtered

```{r}
hNPCs_Filtered <- subset(hNPCs_Filtered, subset = nCount_RNA > 250 & nCount_RNA < 20000 & nFeature_RNA > 200 & nFeature_RNA < 5000 & mitoPercent < 20)

# get the sense of no. of cells and features in filtered seurat object
hNPCs_Filtered

# lets check the no. of cells in each conditions post_filtering
table(hNPCs_Filtered$Condition)

# save the filtered object
saveRDS(hNPCs_Filtered, file = "hNPCs_Filtered.rds")

# View metadata
View(hNPCs_Filtered@meta.data)
```


# Visualize post filtering QC metrics as a violin plot
```{r}

# pdf('ViolinPlot_postfitering.pdf',height=6,width=12)
# VlnPlot(hNPCs_Filtered, features = c("nFeature_RNA", "nCount_RNA", "mitoPercent", "riboPercent"), ncol = 4, pt.size=.1)
# dev.off()
# 
# pdf('ViolinPlot_withPoint_postfitering3.pdf',height=4,width=12)
# VlnPlot(hNPCs_Filtered, features = c("nFeature_RNA", "nCount_RNA", "mitoPercent", "riboPercent"), ncol = 4, pt.size=.01, group.by = "Condition",  cols = c(invitro = "hotpink3", invivoP60 = "#abcc94", invivoP90 = "darkgoldenrod3")) + labs(y = "Your Y-axis Title")
# dev.off()
# 
# pdf('ViolinPlot_post_fitering2.pdf',height=6,width=12)
# VlnPlot(hNPCs_Filtered, features = c("nFeature_RNA", "nCount_RNA", "mitoPercent", "riboPercent"), ncol = 4, pt.size=0, group.by = "Condition")
# dev.off()

pdf('FeatureScatterPlot_postfiltering.pdf',height=5,width=18)
plot1 <- FeatureScatter(hNPCs_Filtered, feature1 = "nCount_RNA", feature2 = "mitoPercent", group.by = "Condition")
plot2 <- FeatureScatter(hNPCs_Filtered, feature1 = "nCount_RNA", feature2 = "riboPercent", group.by = "Condition")
plot3 <- FeatureScatter(hNPCs_Filtered, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "Condition") + geom_smooth(method = lm)
plot1 + plot2 + plot3
dev.off()

# Load patchwork package
library(patchwork)

# Open a PDF file to save the violin plot
tiff('ViolinPlot_withPoint_postfitering_2.tiff',height=3.2,width=11, res=300, units='in')
# Generate individual violin plots
p1 <- VlnPlot(hNPCs_Filtered, features = "nFeature_RNA", pt.size = 0.001, group.by = "Condition", cols = c(invitro = "hotpink3", invivoP60 = "#abcc94", invivoP90 = "darkgoldenrod3")) + 
  NoLegend() + 
  ylab(expression(italic(N)[Genes]))+
  theme(plot.title = element_blank(), axis.title.x = element_blank())

p2 <- VlnPlot(hNPCs_Filtered, features = "nCount_RNA", pt.size = 0.001, group.by = "Condition", cols = c(invitro = "hotpink3", invivoP60 = "#abcc94", invivoP90 = "darkgoldenrod3")) + 
  NoLegend()  + 
  ylab(expression(italic(N)[UMIs])) + 
  theme(plot.title = element_blank(), axis.title.x = element_blank())

p3 <- VlnPlot(hNPCs_Filtered, features = "mitoPercent", pt.size = 0.001, group.by = "Condition", cols = c(invitro = "hotpink3", invivoP60 = "#abcc94", invivoP90 = "darkgoldenrod3")) + 
  NoLegend() + 
  labs(y = "Mito Percent") + 
  theme(plot.title = element_blank(), axis.title.x = element_blank())

p4 <- VlnPlot(hNPCs_Filtered, features = "riboPercent", pt.size = 0.001, group.by = "Condition", cols = c(invitro = "hotpink3", invivoP60 = "#abcc94", invivoP90 = "darkgoldenrod3")) + 
  NoLegend() + 
  labs(y = "Ribo Percent") + 
  theme(plot.title = element_blank(), axis.title.x = element_blank())

# Combine the plots into one with patchwork and arrange them in a grid
(p1 | p2 | p3 | p4) + plot_layout(ncol = 4)

# Close the PDF device
dev.off()


tiff('ViolinPlot_withPoint_postfitering_4.tiff',height=3.2,width=11, res=300, units='in')
# Generate individual violin plots
p1 <- VlnPlot(hNPCs_Filtered, features = "nFeature_RNA", pt.size = 0.001, group.by = "Condition") + 
  NoLegend() + 
  ylab(expression(italic(N)[Genes]))+
  theme(plot.title = element_blank(), axis.title.x = element_blank())

p2 <- VlnPlot(hNPCs_Filtered, features = "nCount_RNA", pt.size = 0.001, group.by = "Condition") + 
  NoLegend()  + 
  ylab(expression(italic(N)[UMIs])) + 
  theme(plot.title = element_blank(), axis.title.x = element_blank())

p3 <- VlnPlot(hNPCs_Filtered, features = "mitoPercent", pt.size = 0.001, group.by = "Condition") + 
  NoLegend() + 
  labs(y = "Mito Percent") + 
  theme(plot.title = element_blank(), axis.title.x = element_blank())

p4 <- VlnPlot(hNPCs_Filtered, features = "riboPercent", pt.size = 0.001, group.by = "Condition") + 
  NoLegend() + 
  labs(y = "Ribo Percent") + 
  theme(plot.title = element_blank(), axis.title.x = element_blank())

# Combine the plots into one with patchwork and arrange them in a grid
(p1 | p2 | p3 | p4) + plot_layout(ncol = 4)

# Close the PDF device
dev.off()

```

# use table function to get the number of cells in each Sample as a dataframe

```{r}

df <- as.data.frame(rev(table(hNPCs_Filtered$Condition)))
colnames(df) <- c('Condition', 'n_cells')

# Define custom colors for each condition
cols <- c(invitro = "hotpink3", invivoP60 = "#abcc94", invivoP90 = "darkgoldenrod3")

# bar plot of the number of cells in each sample
p7 <- ggplot(df, aes(y=n_cells, x=reorder(Condition, -n_cells), fill=Condition)) +
  geom_bar(stat='identity') +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = cols) +  
  NoLegend() + RotatedAxis() +
  ylab(expression(italic(N)[cells])) + xlab('Condition') +
  ggtitle(paste('Post Filter:', sum(df$n_cells))) +
  theme(
    panel.grid.minor=element_blank(),
    panel.grid.major.y=element_line(colour="lightgray", size=0.5),
  )

png('Total_cells_postfiltering.png', width=3, height=4, res=200, units='in')
print(p7)
dev.off()

```

# Create metadata dataframe
Visualize the correlation between genes detected and number of UMIs and determine whether strong presence of cells with low numbers of genes/UMIs

```{r}
View(hNPCs_Filtered@meta.data)
metadata_F <- hNPCs_Filtered@meta.data

p8 <- metadata_F %>% 
  ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=mitoPercent)) + 
  geom_point() + 
  scale_colour_gradient(low = "gray90", high = "black") +
  stat_smooth(method=lm) +
  scale_x_log10() + 
  scale_y_log10() + 
  theme_classic() +
  geom_vline(xintercept = 500) +
  geom_hline(yintercept = 300) +
  facet_wrap(~Condition)

png('PostFilter_Corellation_plot.png', width=10, height=4, res=200, units='in')
print(p8)
dev.off()

# # save filtered object
saveRDS(hNPCs_Filtered, file = "hNPCs_Filtered.rds")
```

#
# load and assign data to new object
```{r}

hNPCs_Filtered <- readRDS("~/Desktop/08-10-2023_hNPC_scRNA_seq_analysis/08102023_hNPC_Filtered/hNPCs_Filtered.rds")

table(hNPCs_Filtered$Sample_ID)
# assign data to new object
hNPCs_Scaled <- hNPCs_Filtered
```

##  create new directory to save the figures
```{r}

dir.create("10042023_hNPC_Scale_data")
setwd("10042023_hNPC_Scale_data")

```

# Check whether data have batch effect or not? 

# 1. Normalize data ----------
```{r}
hNPCs_Scaled <- NormalizeData(hNPCs_Filtered, normalization.method = "LogNormalize", scale.factor = 10000)

```


# 2. Identify and plot highly variable features --------------
```{r}

hNPCs_Scaled <- FindVariableFeatures(hNPCs_Scaled, selection.method = "vst", nfeatures = 3000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(hNPCs_Scaled), 10)

# plot variable features with and without labels
pdf('VariableFeatures.pdf', height=4,width=12)
plot1 <- VariableFeaturePlot(hNPCs_Scaled)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = T)
plot1 + plot2
dev.off()

str(hNPCs_Scaled)
```

# 3. Scale the data -------------
```{r}
all.genes <- rownames(hNPCs_Scaled)
hNPCs_Scaled <- ScaleData(hNPCs_Scaled, features = all.genes)
saveRDS(hNPCs_Scaled, "hNPCs_Scaled.rds")
```


# 4. Run PCA
```{r}
hNPCs_Scaled_PCA <- RunPCA(hNPCs_Scaled, features = VariableFeatures(object = hNPCs_Scaled))

# visualize PCA results
print(hNPCs_Scaled_PCA[["pca"]], dims = 1:5, nfeatures = 5)

# plot the top genes contributing to the first 3 PCs
plot4 <- VizDimLoadings(hNPCs_Scaled_PCA, dims = 1:6, reduction = "pca", ncol = 3)
png('PCA_VizDimLoadings.png', height=12, width=10, res = 200, units = 'in')
print(plot4)
dev.off()

# PCA scatter plot colored by sampleID

plot5 <- DimPlot(hNPCs_Scaled_PCA, reduction = "pca", group.by = 'Condition')
png('PCA_ScatterPlot.png', height=6, width=7, res = 200, units = 'in')
print(plot5)
dev.off()

# Determine dimensionality of the data i.e standard deviation of each PCA in a Elbow plot
png('PCA_ElbowPlot.png',height=3,width=6, res = 200, units = 'in')
ElbowPlot(hNPCs_Scaled_PCA, ndims = 50)
dev.off()

# plot pca map
png('PCA_map.png', width=10, height=10, res=300, units='in')
DoHeatmap(hNPCs_Scaled_PCA, dims = 1:15, cells = 500, balanced = TRUE)
dev.off()

### save the data as RDS file
saveRDS(hNPCs_Scaled_PCA, file = "hNPCs_Scaled_PCA.rds")
```


# perform non-linear dimensionality reduction for visualization--------------
```{r}

hNPCs_rat_UMAP <- FindNeighbors(object = hNPCs_Scaled_PCA, dims = 1:30)
hNPCs_rat_UMAP <- FindClusters(object = hNPCs_rat_UMAP, resolution = c(0.1, 0.2, 0.3, 0.4, 0.5))
hNPCs_rat_UMAP <- RunUMAP(object = hNPCs_rat_UMAP, dims = 1:30)
hNPCs_rat_UMAP <- RunTSNE(object = hNPCs_rat_UMAP, dims = 1:30)

# View metadata
View(hNPCs_rat_UMAP@meta.data)
```

# create new directory to save the figures
```{r}
Idents(hNPCs_rat_UMAP) <- "RNA_snn_res.0.1"
png('UMAP_tSNE_Plots.png',height=6,width=8, res = 200, units = 'in')
DimPlot(object = hNPCs_rat_UMAP, reduction = "tsne",label=TRUE)
DimPlot(hNPCs_rat_UMAP, reduction = "umap",label=TRUE)
dev.off()

png('UMAP_Plot.png',height=6,width=7, res = 200, units = 'in')
DimPlot(hNPCs_rat_UMAP, reduction = "umap", label = T)
dev.off()

png('UMAP_Plot_grouped.png',height=6,width=7, res = 200, units = 'in')
DimPlot(hNPCs_rat_UMAP, reduction = "umap", group.by = "Condition", label = T)
dev.off()

png('UMAP_Plots_split_group.png',height=4,width=10, res = 200, units = 'in')
DimPlot(hNPCs_rat_UMAP, reduction = "umap", split.by = "Condition", label = T)
dev.off()

saveRDS(hNPCs_rat_UMAP, file = "hNPCs_rat_UMAP.rds")

```

# Batch correction:

```{r}
dir.create("Batch_correction")
setwd("Batch_correction")

```


# Integrate the data by CCA____________________
1. set working directory
2. load the data
```{r}

hNPCs_rat_integ <- readRDS("~/Desktop/08-10-2023_hNPC_scRNA_seq_analysis/08102023_hNPC_Filtered/hNPCs_Filtered.rds")

# View metadata
View(hNPCs_rat_integ@meta.data)

rm(hNPCs_rat_UMAP, metadata, metadata_F)
```

# split the dataset into a list of seurat objects (invitro, invivoP60, invivoP90)
```{r}
hNPCs_rat_list <- SplitObject(hNPCs_Filtered, split.by = "Condition")

```


# normalize and identify variable features for each dataset independently
```{r}
hNPCs_rat_list <- lapply(X = hNPCs_rat_list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 3000)
})

```

# Perform integration______________
1. select features that are repeatedly variable across datasets for integration

```{r}
features <- SelectIntegrationFeatures(object.list = hNPCs_rat_list)

```

2. Find integration anchor
```{r}
hNPCs_rat_anchors <- FindIntegrationAnchors(object.list = hNPCs_rat_list, anchor.features = features)

```

3. creates an 'integrated' data assay
```{r}

hNPC_rat_combined <- IntegrateData(anchorset = hNPCs_rat_anchors)

```


# Save integrated seurat object 
```{r}

saveRDS(hNPC_rat_combined, file="hNPC_rat_integ_CCA.rds")

# remove objects that are not required
rm(hNPCs_rat_anchors, hNPCs_rat_list, hNPCs_Scaled, hNPCs_Scaled_PCA)

```


## Perform downstream analysis on the corrected data note that the original unmodified data still resides in the 'RNA' assay, set Defult assay  "integrated"

```{r}
DefaultAssay(hNPC_rat_combined) <- "integrated"
```

# Scale the data and PCA
```{r}
hNPC_rat_combined <- ScaleData(hNPC_rat_combined, verbose = FALSE)
hNPC_rat_combined <- RunPCA(hNPC_rat_combined, npcs = 30, verbose = FALSE)
```

# Plot PCA
```{r}
png('PCA_ScatterPlot.png', height=3, width=7, res = 200, units = 'in')
PCAPlot(hNPC_rat_combined,
        split.by = "Condition")  
dev.off()

png('PCA_ElbowPlot_Batch.png',height=3,width=6, res = 200, units = 'in')
ElbowPlot(hNPC_rat_combined, ndims = 30)
dev.off()

png('PCA_Heatmap_Batch.png', width=10, height=10, res=300, units='in')
DoHeatmap(hNPC_rat_combined, dims = 1:15, cells = 500, balanced = TRUE)
dev.off()

```

# Run the standard workflow for visualization and clustering
```{r}

hNPC_rat_combined <- FindNeighbors(hNPC_rat_combined, reduction = "pca", dims = 1:30)
hNPC_rat_combined <- FindClusters(hNPC_rat_combined, resolution =  c(0.1, 0.2, 0.3, 0.4, 0.5))
hNPC_rat_combined <- RunUMAP(hNPC_rat_combined, reduction = "pca", dims = 1:30)
hNPC_rat_combined <- RunTSNE(object = hNPC_rat_combined, reduction = "pca", dims = 1:30)

# View metadata
View(hNPC_rat_combined@meta.data)
```


# Assign identity of clusters
```{r}

Idents(object = hNPC_rat_combined) <- "integrated_snn_res.0.5"

#check number of cells in each cluster
table(Idents(hNPC_rat_combined))
```


# save figure in new directory
```{r}

dir.create('integ_data_fig')
setwd("integ_data_fig")

```


#check the Umap plot after CCA Batch Correction
```{r}

umap_theme <- theme(
  axis.line = element_line(),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_line(),
  axis.title.x = element_text(size = 12),
  axis.title.y = element_text(size = 12),
  panel.background = element_blank(),
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  plot.background = element_blank(),
  plot.title = element_text(size = 12)
)

png('hNPC_rat_umap_tsne_grouped.png', height=4, width=10, res = 300, units = 'in')
p10 <- DimPlot(hNPC_rat_combined, reduction = "tsne", group.by = "Condition", pt.size = 0.1)
p11 <- DimPlot(hNPC_rat_combined, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.1) + umap_theme
p10 + p11
dev.off()


png('CCA_umap_splitted.png', height=4, width=10, res = 200, units = 'in')
DimPlot(hNPC_rat_combined, reduction = 'umap', split.by = "Condition", label = T) + NoLegend()
dev.off()


png('CCA_umap_grouped.png', height=5, width=6, res = 200, units = 'in')
DimPlot(hNPC_rat_combined, reduction = 'umap', group.by = "Condition", label = T) 
dev.off()

```


# Save integrated seurat object 
```{r}

saveRDS(hNPC_rat_combined, file="hNPC_rat_combined_UMAP.rds")

```


# identify clusters on the basis of markers genes
```{r}
genes=c("HLA-DMA", "HLA-DR","HLA-G", "HLA-C", "HLA-A", "HLA-B","HLA-E","HLA-DRB1", "HLA-DRA", "HLA-F","TMEM119", "CX3CR1", 'P2RY12', "CSF1R", "CD99", "TYROBP","ITGAM","NES", "SOX2", "CD44", "HOPX", "SOX9", "S100B", "VIM", "GFAP", "AQP4", "SLC1A2", "SLC1A3", "OLIG1", "OLIG2", "MOBP", "MBP", "MOG", "PDFGRA", "CSPG4","GAD2", "MAP2", "SNAP25", "SYT1", "PLP1", "PLP4", "GRM6", "PMP2", "DTNA", "RHO", "RCVRN", "GNAT1","ROM1", "ARR","GNAT2",  "OPN1SW", "RBPMS", "RLBP1","GLUL", "VSX2", "VSX1","GNK1","GSG1","SNGC", "GNGT2", "CALB1", "CAPN14", "SCA1", "CD117", "CD45RA", "CD14", "CD68", "CD163", "CD45", "MANF", "CDNF","GDNF", "BDNF", "NFE2L2", "FGF1", "FGF2", "MYDGF","IGF", "MANFR", "ERN1","KDELR1" ,"KDELR2", "NTRK1", "NTRK2")


png('MarkerGenes_Plots.png', height=150,width=10, res = 200, units = 'in')
FeaturePlot(hNPC_rat_combined, features = genes, max.cutoff = 98, min.cutoff = "q10", split.by = "Condition", cols = viridis(256), pt.size = 0.8, reduction = "umap") & umap_theme
dev.off()


png('VlnMarkerGenes_Plots.png', height=70,width=30, res = 200, units = 'in')
VlnPlot(hNPC_rat_combined, features = genes, pt.size=0)
dev.off()

```

# Plot heatmap (assay = integrated)
```{r}

png('Markergenes_map_integ.png', width=10, height=10, res=300, units='in')
DoHeatmap(hNPC_rat_combined, features = genes, assay = "integrated") + NoLegend()
dev.off()

png('Markergenes_map_integ_Vir.png', width=10, height=10, res=300, units='in')
DoHeatmap(hNPC_rat_combined, features = genes, assay = "integrated" , size = 5.5) + NoLegend() + scale_fill_gradientn(colors=viridis(256)) 
dev.off()
```

# Plot heatmap (assay=RNA)
```{r}

png('Markergenes_map_RNA.png', width=10, height=10, res=300, units='in')
DoHeatmap(hNPC_rat_combined, features = genes, assay = "RNA") + NoLegend()
dev.off()

```


# Find conserved marker for 1, 3 and 12 clusters
```{r}
## set DefaultAssay to RNA
DefaultAssay(hNPC_rat_combined) <- "RNA"

## C1
C1.markers <- FindConservedMarkers(hNPC_rat_combined, ident.1 = 1, grouping.var = "Condition", verbose = FALSE)
head(C1.markers, 15)

## load annotation file
annotations <- read.csv("annotations.csv")

# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C1.markers <- C1.markers %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() 

# Write all results to file
write.csv(C1.markers,"C1_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)

## C3
C3.markers <- FindConservedMarkers(hNPC_rat_combined, ident.1 = 3, grouping.var = "Condition", verbose = FALSE)
head(C3.markers, 15)

C3.markers <- C3.markers %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() 

# Write all results to file
write.csv(C3.markers,"C3_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)

## C12
C12.markers <- FindConservedMarkers(hNPC_rat_combined, ident.1 = 12, grouping.var = "Condition", verbose = FALSE)
head(C12.markers, 15)
C12.markers <- C12.markers %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() 

# Write all results to file
write.csv(C12.markers,"C12_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


```

# Rename suerat cluster__________
```{r}
cluster_annotations <- list(
  '0' = 'hNPC',
  '1' = 'Microglia1',
  '2' = 'Bipolar1',
  '3' = 'Microglia2',
  '4' = 'hNPC',
  '5' = 'Müller glia',
  '6' = 'RGC1', 
  '7' = 'hNPC',
  '8' = 'hNPC',
  '9' = 'Bipolar2',
  '10' = 'PR',
  '11' = 'RGC2',
  '12' = 'Microglia3'
  )
```


# Add CellType to seurat metadata__________
```{r}

hNPC_rat_combined$CellType <- unlist(cluster_annotations[hNPC_rat_combined$integrated_snn_res.0.1])
hNPC_rat_combined$CellType_cluster <- paste0(hNPC_rat_combined$CellType, '-', hNPC_rat_combined$integrated_snn_res.0.1)

# Check the metadata for renamed cluster 
View(hNPC_rat_combined@meta.data)


```

# Plot UMAP for integrated dataset__________
```{r}

umap_theme1 <- theme(
  axis.line = element_line(),
  axis.text.x = element_text(size = 15),
  axis.text.y = element_text(size = 15),
  axis.ticks = element_line(),
  axis.title.x = element_text(size = 15),
  axis.title.y = element_text(size = 15),
  panel.background = element_blank(),
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  plot.background = element_blank(),
  plot.title = element_text(size = 15)
)

png('umap_celltypes_grouped.png', width=8, height=7, res=200, units='in')
DimPlot(hNPC_rat_combined, label = TRUE, group.by = "CellType") & umap_theme1
dev.off()

```

# Plot map (assay = integrated)
```{r}

genes=c("HLA-A", "HLA-B","HLA-C","NES", "SOX2", "PLP1", "PLP4", "GRM6", "PMP2", "DTNA","VIM", "CD44", "HOPX", "SOX9", "S100B","GFAP", "AQP4", "SLC1A2", "EGFR","PDFGRA","CSPG4", "OLIG1", "OLIG2", "MOBP", "MBP", "MOG", "CXCR4", "MAP2", "GAD2","SNAP25", "SYT1",  "RHO", "RCVRN", "GNAT1","ROM1", "ARR","GNAT2",  "OPN1SW", "RBPMS", "RLBP1","GLUL","VSX1", "VSX2", "GNK1","GSG1","SNGC", "GNGT2", "CALB1", "CAPN14", "SCA1", "CD117", "CD45RA", "CD14", "CD68", "CD163", "CD45", "CX3CR1",  "CSF1R","TYROBP")

png('Markergenes_map_integ.png', width=10, height=10, res=300, units='in')
DoHeatmap(hNPC_rat_combined, features = genes, assay = "integrated", group.by = "integrated_snn_res.0.1") + NoLegend()
dev.off()

png('Markergenes_map_integ_Vir.png', width=10, height=10, res=300, units='in')
DoHeatmap(hNPC_rat_combined, features = genes, assay = "integrated" , size = 5.5) + NoLegend() + scale_fill_gradientn(colors=viridis(256)) 
dev.off()

png('Markergenes_map_integ.png', width=10, height=10, res=300, units='in')
DotPlot(hNPC_rat_combined, features = genes, assay = "integrated", group.by = "integrated_snn_res.0.1") + NoLegend()
dev.off()

png('Dotplot_Markergenes_integ.png', width=18, height=5, res=300, units='in')
DotPlot(hNPC_rat_combined, features = genes, assay = "RNA") + RotatedAxis()
dev.off()

png('Dotplot_Markergenes_integ1.png', width=18, height=5, res=300, units='in')
DotPlot(hNPC_rat_combined, features = genes, assay = "RNA", group.by = "CellType") + RotatedAxis()
dev.off()

# Save the object
saveRDS(hNPC_rat_combined, file="hNPC_rat_combined_clusterID.rds")
```

# Subset the hNPC Cluster________
# set directories__________
```{r}

dir.create('subset_hNPC_cluster')
setwd('subset_hNPC_cluster')

```


# change DefaultAssay and Idents__________
```{r}

DefaultAssay(hNPC_rat_combined) <- "RNA"
Idents(hNPC_rat_combined) <- "CellType"

```

# Plot UMAP for integrated dataset__________
```{r}
png('umap_celltypes_grouped1.png', width=8, height=7, res=200, units='in')
DimPlot(hNPC_rat_combined, label = TRUE) & umap_theme1
dev.off()

```

# subset hNPC_Cluster
```{r}

hNPC_cluster <- subset(hNPC_rat_combined, idents = "hNPC")
rm(hNPC_rat_combined)

# Save the object as RDS file
saveRDS(hNPC_cluster, file = "subset_hNPCs_Cluster.rds")

```

# visualization
```{r}

pdf('ViolinPlot_withPoint.pdf', height=3.5,width=10)
VlnPlot(hNPC_cluster, features = c("nFeature_RNA", "nCount_RNA", "mitoPercent", "riboPercent"), ncol = 4, pt.size=.2, group.by = "Condition")
dev.off()

png('hNPC_UMAP_grouped.png', height=5,width=6, res=200, units='in')
DimPlot(hNPC_cluster, reduction = 'umap', group.by = 'Condition')
dev.off()

png('hNPC_UMAP_Splitted.png', height=5,width=12, res=200, units='in')
DimPlot(hNPC_cluster, reduction = 'umap', pt.size = 0.5 , split.by = 'Condition')
dev.off()

```


# check the total no. of features and cells in hNPCs cluster or different condition
```{r}
hNPC_cluster
table(hNPC_cluster$Condition)
```

# Extract number of cells per cluster per orig.ident 
```{r}

n_cells <- FetchData(hNPC_cluster, vars = c("ident", "Condition")) %>%
  dplyr::count(ident, Condition) %>%
  tidyr::spread(ident, n)

# save the cell no. in CSV file
write.csv(n_cells, file = "Cells_per_cluster.csv")

n_cells

rm(n_cells)

```


# Perform integration and following steps again on subset hNPC cluster_____

1. Split the dataset into a list of seurat objects (invitro, invivoP60, invivoP90)
```{r}

hNPC.list <- SplitObject(hNPC_cluster, split.by = "Condition")

```


# Normalize and identify variable features for each dataset independently__________
```{r}
hNPC.list <- lapply(X = hNPC.list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 3000)
})
```


# Perform integration______________
# select features that are repeatedly variable across datasets for integration
```{r}

features <- SelectIntegrationFeatures(object.list = hNPC.list)

```


# Find integration anchor
```{r}
hNPC.anchors <- FindIntegrationAnchors(object.list = hNPC.list, anchor.features = features)

```

# creates an 'integrated' data assay
```{r}

hNPC_integ <- IntegrateData(anchorset = hNPC.anchors)

# save object
saveRDS(hNPC_integ, "hNPC_integ.rds")

```


# Scale the data and run PCA_________
```{r}
## set default assay to integrated
DefaultAssay(hNPC_integ) <- "integrated"
hNPC_integ <- ScaleData(hNPC_integ, verbose = FALSE)
hNPC_integ <- RunPCA(hNPC_integ, npcs = 30, verbose = FALSE)

```


# save figure in new directory
```{r}

dir.create('hNPC_integ_data_fig')
setwd("hNPC_integ_data_fig")

```


# Visualize the PCA plot
```{r}

png('PCA_map.png',height=10,width=10, res = 200, units = 'in')
DoHeatmap(hNPC_integ, dims = 1:15, cells = 500, balanced = TRUE)
dev.off()


# PCA scatter plot colored by sampleID
png('PCA_ScatterPlot.png', height=6, width=7, res = 200, units = 'in')
DimPlot(hNPC_integ, reduction = "pca", group.by = 'Condition')
dev.off()



png('PCA_ElbowPlot.png', height=3, width=6, res = 200, units = 'in')
ElbowPlot(hNPC_integ)
dev.off()

```


# Run the standard workflow for clustering_________
```{r}
hNPC_integ <- FindNeighbors(hNPC_integ, reduction = "pca", dims = 1:20)
hNPC_integ <- FindClusters(hNPC_integ, resolution =  c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9))
hNPC_integ <- RunUMAP(hNPC_integ, reduction = "pca", dims = 1:20)
hNPC_integ <- RunTSNE(hNPC_integ, reduction = "pca", dims = 1:20)


View(hNPC_integ@meta.data)
```

# check the Umap plot after CCA Batch Correction
```{r}

View(hNPC_integ@meta.data)

Idents(hNPC_integ) <- "integrated_snn_res.0.4"
pdf('hNPC_umap_tsne_integrated_snn_res.0.4.pdf',height=4,width=10)
p14 <- DimPlot(hNPC_integ, reduction = "tsne", group.by = "Condition", pt.size = 0.8)
p15 <- DimPlot(hNPC_integ, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.8)
p14 + p15
dev.off()


Idents(hNPC_integ) <- "integrated_snn_res.0.5"
pdf('hNPC_umap_tsne_integrated_snn_res.0.5.pdf',height=4,width=10)
p14 <- DimPlot(hNPC_integ, reduction = "tsne", group.by = "Condition", pt.size = 0.8)
p15 <- DimPlot(hNPC_integ, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.8)
p14 + p15
dev.off()

Idents(hNPC_combined) <- "integrated_snn_res.0.6"
pdf('hNPC_umap_tsne_integrated_snn_res.0.6.pdf',height=4,width=10)
p14 <- DimPlot(hNPC_integ, reduction = "tsne", group.by = "Condition", pt.size = 0.8)
p15 <- DimPlot(hNPC_integ, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.8)
p14 + p15
dev.off()

Idents(hNPC_combined) <- "integrated_snn_res.0.7"
pdf('hNPC_umap_tsne_integrated_snn_res.0.7.pdf',height=4,width=10)
p14 <- DimPlot(hNPC_integ, reduction = "tsne", group.by = "Condition", pt.size = 0.8)
p15 <- DimPlot(hNPC_integ, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.8)
p14 + p15
dev.off()

Idents(hNPC_integ) <- "integrated_snn_res.0.8"
pdf('hNPC_umap_tsne_integrated_snn_res.0.8.pdf',height=4,width=10)
p14 <- DimPlot(hNPC_integ, reduction = "tsne", group.by = "Condition", pt.size = 0.8)
p15 <- DimPlot(hNPC_integ, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.8)
p14 + p15
dev.off()


Idents(hNPC_integ) <- "integrated_snn_res.0.9"
pdf('hNPC_umap_tsne_integrated_snn_res.0.9.pdf',height=4,width=10)
p14 <- DimPlot(hNPC_integ, reduction = "tsne", group.by = "Condition", pt.size = 0.8)
p15 <- DimPlot(hNPC_integ, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.8)
p14 + p15
dev.off()

```

# Marker gene
```{r}
genes=c("NES", "SOX2", "CD44", "GLAST","VIM", "PAX6", "CXCR4", "NEUROD1","MAP2", "NEUN", 'SNAP25', "SYT1", "EGFR", 'PDGFRA', 'CSPG4', "OLIG1", "OLIG2", "MBP", "MOG","MOBP","HES5","SOX9", "HOPX", "PAX3", "GFAP","S100B", "AQP4", "GLUL", "SLC1A2", "CD24", "THY1")

png('Markergenes_heatmap_assay_int.png', width=6, height=8, res=300, units='in')
DoHeatmap(object = hNPC_integ, features = genes, assay = "integrated", label = F,
  size = 5.5) + theme(axis.text.y = element_text(size = 10))
dev.off()

png('Markergenes_heatmap_assay_RNA.png', width=6, height=8, res=300, units='in')
DoHeatmap(object = hNPC_integ, features = genes, assay = "RNA", label = F,
  size = 5.5) + theme(axis.text.y = element_text(size = 10))
dev.off()


# visualize
png('hNPC_umap_splitted2.png',height=4,width=10, res=300, units='in')
DimPlot(hNPC_integ, reduction = 'umap', split.by = "Condition", label = T, pt.size = 0.8) + NoLegend()
dev.off()

png('hNPC_umap_grouped3.png',height=5,width=6, res=300, units='in')
DimPlot(hNPC_integ, reduction = 'umap', group.by = "Condition", label = T, pt.size = 0.5) 
dev.off()


```

# umap
```{r}
umap_theme <- theme(
  axis.line = element_line(),
  axis.text.x = element_text(size = 15),
  axis.text.y = element_text(size = 15),
  axis.ticks = element_line(),
  axis.title.x = element_text(size = 15),
  axis.title.y = element_text(size = 15),
  panel.background = element_blank(),
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  plot.background = element_blank(),
  plot.title = element_text(size = 15)
)


Idents(hNPC_integ) <- "integrated_snn_res.0.6"
png('hNPC_umap_tsne_grouped.png', height=4, width=10, res = 300, units = 'in')
p21 <- DimPlot(hNPC_integ, reduction = "tsne", group.by = "Condition", pt.size = 0.5)
p22 <- DimPlot(hNPC_integ, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5) + umap_theme
p21 + p22
dev.off()

```

# check number of cells in each cluster (integrated_snn_res.0.6)
```{r}

table(hNPC_integ$integrated_snn_res.0.6)

# save data in csv file
n_cells <- FetchData(hNPC_integ, vars = c("ident", "Condition")) %>%
  dplyr::count(ident, Condition) %>%
  tidyr::spread(ident, n)

n_cells1

write.csv(n_cells, file = "Cells_per_cluster_integrated_snn_res.0.6.csv")

rm(n_cells)

```

# save RDS
```{r}
saveRDS(hNPC_integ, "hNPC_integ_Cluster.rds")
```

