---
title: "hNPC_scRNAseq_analysis"
author: "SabaShahin"
date: "06/06/2023"
output: html_document
    
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


load required libraries
```{r}
suppressWarnings(library(dplyr))
suppressWarnings(library(Seurat))
suppressWarnings(library(tidyverse))
suppressWarnings(library(Matrix))
suppressWarnings(library(ggplot2))
suppressWarnings(library(cowplot))
suppressWarnings(library(units))
suppressWarnings(library(umap))
suppressWarnings(library(cluster))
suppressWarnings(library(tsne))
suppressWarnings(library(viridis))
suppressWarnings(library(patchwork))
suppressWarnings(library(pmap))
suppressWarnings(library(irlba))
suppressWarnings(library(beepr))
suppressWarnings(library(gridExtra))
suppressWarnings(library(Rcpp))
suppressWarnings(library(png))
suppressWarnings(library(harmony))
```


##################################################### Group_DEG  ################################################################################

## load data or assign data to new object
```{r}
setwd("~/Desktop/08-10-2023_hNPC_scRNA_seq_analysis/subset_hNPC_cluster/hNPC_integ_data_fig")
hNPC <- readRDS("hNPC_integ_Cluster.rds")

```

## create new directory for cluster analysis

```{r}
dir.create("hNPC_GrpDEG_analysis")
setwd("hNPC_GrpDEG_analysis")
```

## filter genes with zero count

```{r}
DefaultAssay(hNPC) <- "RNA"
hNPC_cts <- GetAssayData(hNPC , slot='counts')

## filter the genes
hNPC_cts <- rownames(hNPC_cts)[rowSums(hNPC_cts) >= 10]

hNPC <- hNPC[hNPC_cts,]


# Normalize data ----------
hNPC <- NormalizeData(hNPC, normalization.method = "LogNormalize", scale.factor = 10000)

# 2. Find VariableFeatures
hNPC <- FindVariableFeatures(hNPC, selection.method = "vst", nfeatures = 3000)
# Scale the data ----------
all.genes <- rownames(hNPC)
hNPC <- ScaleData(hNPC, features = all.genes)

## save the filtered object
saveRDS(hNPC, "hNPC_fil.rds")

```


# Perform DEG using MAST:
```{r}
Idents(hNPC) <- "Condition"

invivoP60_invitro_DEG <- FindMarkers(hNPC, ident.1 = "invivoP60", ident.2 = "invitro",  verbose = FALSE, test.use = "MAST", logfc.threshold = 0.10, min.pct = 0.10)
write.csv(invivoP60_invitro_DEG,"invivoP60_invitro_DEG.csv", quote=F)

invivoP90_invitro_DEG <- FindMarkers(hNPC, ident.1 = "invivoP90", ident.2 = "invitro",  verbose = FALSE, test.use = "MAST", logfc.threshold = 0.10, min.pct = 0.10)
write.csv(invivoP90_invitro_DEG,"invivoP90_invitro_DEG.csv", quote=F)


invivoP90_invivoP60_DEG <- FindMarkers(hNPC, ident.1 = "invivoP90", ident.2 = "invivoP60",  verbose = FALSE, test.use = "MAST", logfc.threshold = 0.10, min.pct = 0.10)
write.csv(invivoP90_invivoP60_DEG,"invivoP90_invivoP60_DEG.csv", quote=F)


```


## create new directory for DEG excluding ribosomal genes
```{r}
dir.create("DEGs_Excluding_RibosomalGenes")
setwd("DEGs_Excluding_RibosomalGenes")
```

## get a list of non-ribosomal genes.
```{r}
genes.use <- grep(pattern = "^RP[SL][[:digit:]]|^RP[[:digit:]]|^RPSA",
                  rownames(hNPC),
                  value=TRUE, invert=TRUE) #get list of non-ribosomal genes 
```



# Perform DEG using MAST (excluding ribosomal genes):
```{r}

invivoP60_invitro_DEG_Ex_Ribo <- FindMarkers(hNPC, ident.1 = "invivoP60", ident.2 = "invitro", logfc.threshold = 0.10, min.pct = 0.10, verbose = FALSE, features = genes.use, test.use = "MAST")
write.csv(invivoP60_invitro_DEG_Ex_Ribo,"invivoP60_invitro_DEG_Ex_Ribo.csv")

invivoP90_invitro_DEG_Ex_Ribo <- FindMarkers(hNPC, ident.1 = "invivoP90", ident.2 = "invitro", logfc.threshold = 0.10, min.pct = 0.10, verbose = FALSE, features = genes.use, test.use = "MAST")
write.csv(invivoP90_invitro_DEG_Ex_Ribo,"invivoP90_invitro_DEG_Ex_Ribo.csv")

invivoP90_invivoP60_DEG_Ex_Ribo <- FindMarkers(hNPC, ident.1 = "invivoP90", ident.2 = "invivoP60", logfc.threshold = 0.10, min.pct = 0.10, verbose = FALSE, features = genes.use, test.use = "MAST")
write.csv(invivoP90_invivoP60_DEG_Ex_Ribo,"invivoP90_invivoP60_Ex_Ribo.csv")


```

##################################################### Cluster_markers ################################################################################
## set working directory
```{r}
setwd("~/Desktop/08-10-2023_hNPC_scRNA_seq_analysis/subset_hNPC_cluster/hNPC_integ_data_fig")

dir.create("Conserved_and_All_markers")
setwd("Conserved_and_All_markers")

View(hNPC@meta.data)
```

## Find AllMarkers to identify the cell-type
```{r}
# Find markers for every cluster compared to all remaining cells, report only the positive ones
Idents(hNPC) <- "integrated_snn_res.0.6"
DefaultAssay(hNPC) <- "RNA"

DimPlot(hNPC, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5) 

Cluster_markers <- FindAllMarkers(object = hNPC, 
                          only.pos = TRUE,
                          logfc.threshold = 0.25) 
write.csv(Cluster_markers, "Cluster_markers.csv")



## Cluster0 Conserved markers

C0_hNPC <- FindConservedMarkers(hNPC, ident.1 = 0, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C0_hNPC_res <- C0_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C0_hNPC_res,"C0_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster1 Conserved markers

C1_hNPC <- FindConservedMarkers(hNPC, ident.1 = 1, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C1_hNPC_res <- C1_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C1_hNPC_res,"C1_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster2 Conserved markers

C2_hNPC <- FindConservedMarkers(hNPC, ident.1 = 2, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C2_hNPC_res <- C2_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name"))%>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C2_hNPC_res,"C2_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster3 Conserved markers

C3_hNPC <- FindConservedMarkers(hNPC, ident.1 = 3, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C3_hNPC_res <- C3_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C3_hNPC_res,"C3_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster4 Conserved markers

C4_hNPC <- FindConservedMarkers(hNPC, ident.1 = 4, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C4_hNPC_res <- C4_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C4_hNPC_res,"C4_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster5 Conserved markers
C5_hNPC <- FindConservedMarkers(hNPC, ident.1 = 5, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C5_hNPC_res <- C5_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C5_hNPC_res,"C5_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster6 Conserved markers

C6_hNPC <- FindConservedMarkers(hNPC, ident.1 = 6, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)

# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C6_hNPC_res <- C6_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C6_hNPC_res,"C6_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster7 Conserved markers

C7_hNPC <- FindConservedMarkers(hNPC, ident.1 = 7, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C7_hNPC_res <- C7_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C7_hNPC_res,"C7_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster8 Conserved markers
C8_hNPC <- FindConservedMarkers(hNPC, ident.1 = 8, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C8_hNPC_res <- C8_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C8_hNPC_res,"C8_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster9 Conserved markers

C9_hNPC <- FindConservedMarkers(hNPC, ident.1 = 9, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C9_hNPC_res <- C9_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() %>%
  arrange(invitro_p_val_adj)

# Write all results to file
write.csv(C9_hNPC_res,"C9_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster10 Conserved markers
C10_hNPC <- FindConservedMarkers(hNPC, ident.1 = 10, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C10_hNPC_res <- C10_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() 

# Write all results to file
write.csv(C10_hNPC_res,"C10_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)

## Cluster10 Conserved markers
C11_hNPC <- FindConservedMarkers(hNPC, ident.1 = 11, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C11_hNPC_res <- C11_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() 

# Write all results to file
write.csv(C11_hNPC_res,"C11_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)

## Cluster12 Conserved markers
C12_hNPC <- FindConservedMarkers(hNPC, ident.1 = 12, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C12_hNPC_res <- C12_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() 

# Write all results to file
write.csv(C12_hNPC_res,"C12_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)


## Cluster13 Conserved markers
C13_hNPC <- FindConservedMarkers(hNPC, ident.1 = 13, grouping.var = "Condition", verbose = FALSE, only.pos = TRUE)


# Turn the DESeq2 results object into a tibble for use with tidyverse functions

C13_hNPC_res <- C13_hNPC %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>% 
                left_join(y = unique(annotations[, c("gene_name", "description")]),
                          by = c("gene" = "gene_name")) %>%
  as_tibble() 

# Write all results to file
write.csv(C13_hNPC_res,"C13_ConservedMarkers.csv", quote = FALSE, row.names = FALSE)

```


